Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-09-14 paper/sections/proof-reconstruction.tex
[ paper ] thm-split added.

The proof-reconstruction approach proposed here consists of a series
of steps similar to the workflow presented by Sultana in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is the proof
assistant, \Agda. We choose \Agda, but another proof assistant with
support of Type Theory and similar features like \Agda
(see a description of \Agda in section
\ref{sec:implementation}) is valid; our results do not depend on
the proof assistant but the support of inductive data types,
termination checking, and type-checking.

Following the diagram in Fig.
\ref{fig:proof-reconstruction-workflow}, the process can begin with
a problem in \CPL that we encode in \TPTP format to be sent as the
input of the \Metis prover.
The proof-search algorithms of \Metis in step No. 3 intend to find
a solution; if they success, \Metis replies a derivation in \TSTP
format of the proof, else \Metis may no stop and running forever or
replies the problem is counter feasible.

@@

In the rest of this section, we provide a formal description using
Type Theory to build definitions and theorems for the functions
necessary to emulate the \Metis' inference rules.

@@

\subsection{Well-structured Recursion}
\label{ssec:well-structured recursion}

To emulate the \Metis' Inference Rules in type theory.
We encounter many functions involved with termination problems, that
means, it is not possible to determine effectively if the function
terminates. That makes impossible to translate in type theory
directly. For that reason, we show a technique that will help us to fix such a issue and guide us how to proceed later.

A recursive function is a \emph{well-structured recursion}, if the
size of the arguments with which it is defined decrease in
size for each call to the function in the definition.
A well-structured function terminates and it can be translated
directly to type theory.

Consider the following example from the \strip inference.

\begin{equation}
\label{eq:unshunt-definition}
\unshunt(φ) =
\begin{cases}
\unshunt((φ₁ ∧ φ₂) ⇒ φ₃)
  & φ := φ₁ ⇒ (φ₂ ⇒ φ₃),\\

\unshunt(φ₁ ⇒ φ₂) ∧ \unshunt(φ₁ ⇒ φ₃)
  & φ := φ₁ ⇒ (φ₂ ∧ φ₃),\\

@@

From the above definition, in the first case of the \unshunt
function, the size of the argument doesn't really decrease.
Actually, it maintains with the same size ((φ₁ ∧ φ₂) ⇒ φ₃ and  (φ₁ ⇒
(φ₂ ⇒ φ₃))). Therefore, a termination checking will fail with \unshunt
function.

To improve it, we will use the technique proposed by Bove and Capettra in
\cite{Bove2005}, which in our case basically consists of
take a general recursive function $f: \sigma \to \tau$.
and an predicate $P: \sigma \to \abbre{Ind}$, where $\abbre{Ind}$
is a type with a structure inductive definition.
Then, we redefine the function as
$f: \sigma \to \abbre{Ind}\ \to \tau$, and perform the recursion
using induction on the structure of the \abbre{Ind} type.

In our case, $\abbre{Ind}$ will be the \abbre{Nat} data type for
natural numbers with constructors \abbre{zero} and \abbre{suc} defined as usual.

Thus, we have function $\unshunt : \PROP \to \Nat \to \PROP$ defined by structural recursion in his second term with type as follows,

\begin{align*}
\label{eq:unshunt-structured}
\unshunt(φ, \zero) &= φ\\
\unshunt(φ, \suc(n)) &=
\begin{cases}
\unshunt((φ₁ ∧ φ₂) ⇒ φ₃, n)
  & φ := φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\unshunt(φ₁ ⇒ φ₂, n) ∧ \unshunt(φ₁ ⇒ φ₃,n)
  & φ := φ₁ ⇒ (φ₂ ∧ φ₃).
\end{cases}
\end{align*}

Finally, we define the complexity measure to handle the second argument for the \unshunt function.

@@

A \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017} or
a rank function presented in \cite{VanDalen1994} is a function
$\PROP \to \Nat$ that indents to associate a natural number for a
proposition φ. The following function is a complexity measure.

\begin{equation*}
\label{eq:rank-definition}
\rank(φ)=
\begin{cases}
0 &\text{for atomic }φ, \\
\rank(φ₁) + c_{¬}  & φ := \neg φ₁, \\
\max(\rank(φ₁),\rank(φ₂)) + c_{\square}
  & φ := φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}
where $c_{¬}$ or $c_{\square}$ is a non-negative integer called the
complexity of the connective.

Nevertheless, we need to define an appropriate complexity measure for the second argument in the $\unshunt$ function, in order to
expect the same behavior of the original function as it was defined at first in Eq.~\ref{eq:unshunt-definition}.

\begin{equation}
\label{eq:rank-definition}
\unshunt_{cm}(φ)=
\begin{cases}
\unshunt_{cm}(φ₃) + 2 & φ := φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\max(\unshunt_{cm}(φ₂),\unshunt_{cm}(φ₃)) + 1
  & φ := φ₁ ⇒ (φ₂ ∧ φ₃),\\
0 &\text{otherwise.}
\end{cases}
\end{equation}

Then, a curated version of \unshunt is presented in Eq.
\ref{eq:unshunt-fixed}.

\begin{equation}
\label{eq:unshunt-fixed}
\unshunt(\varphi) := \unshunt(\varphi, \unshunt_{cm}(\varphi))
\end{equation}

In the following section, we may need define such complexity measure
functions for termination-checking.

@@

\emph{Splitting.}
To prove a goal, we can prove smaller subgoals and combine their
proofs to achieve the final target, a proof of the initial goal.

\begin{theorem}[{\tt thm-$\unshunt^{*}$}] $Γ ⊢ \unshunt^{*}(φ, n)$
for $n\in \Nat$, then $Γ ⊢ φ$.
\label{thm:unshunt}
\end{theorem}

\begin{proof} Use induction on the cases defined by the result of
$\unshunt^{*}$. When $n = 0$, by definition in Eq.
\ref{eq:unshunt-structured} then $Γ ⊢ φ$.
For the first case, we prove $Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)$ from
$Γ ⊢ \unshunt^{*}(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$.
Because $n > 0$, using induction on the second argument and the
theorem $∧⇒$\texttt{-to-}$⇒⇒$ from \cite{AgdaProp}.

\begin{equation*}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
\RightLabel{By definition}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
\RightLabel{\tt thm-$\unshunt^{*}$}
\UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
\RightLabel{\tt ∧⇒-to-⇒⇒}
\UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)}
\end{bprooftree}
\end{equation*}

The second case proves $Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)$ from
 $Γ ⊢ \unshunt^{*}(φ₁ ⇒ (φ₂ ∧ φ₃), n)$ and the
theorem $⇒∧⇒$\texttt{-to-}$⇒∧$ from \cite{AgdaProp}
\begin{equation*}
\scalebox{0.95}{
$\mathcal{D}_1$ := \begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂, n) ∧ \unshunt^{*}(φ₁ ⇒ φ₃, n)$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂, n)$}
\RightLabel{By definition}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂, n-1)$}
\RightLabel{\tt thm-$\unshunt^{*}$}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
\end{bprooftree}
}
\end{equation*}

\begin{equation*}
\mathcal{D}_2 :=
\begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂) ∧ \unshunt^{*}(φ₁ ⇒ φ₃)$}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₃, n)$}
\RightLabel{By definition}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₃, n-1)$}
\RightLabel{\tt thm-$\unshunt^{*}$}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
\end{bprooftree}
\end{equation*}

\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}_1$}
\AxiomC{$\mathcal{D}_2$}
\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
\RightLabel{\tt ⇒∧⇒-to-⇒∧}
\UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
\end{bprooftree}
\end{equation*}
\end{proof}

\vskip 3mm

From now on, we redefine the function $\unshunt$ with the
definition of $\unshunt^{+}$ for no overloading the notation forward.

\begin{theorem}[{\tt thm-\unshunt}] $Γ ⊢ \unshunt(φ)$ then $Γ ⊢ φ$.
\end{theorem}
\begin{proof} It follows from Thm.~\ref{thm:unshunt} using the
\rank function as the second argument.
\end{proof}

The \mysplit function yields a conjunction of subformulas that
implies the proposition φ, the argument of the function. This
function has type $\PROP \to \NN \to PROP$, its definition is
 $\mysplit(φ,0) = φ$, and
for all$n\in \NN$, $n > 0$,

\begin{equation}
\label{eq:split-definition}
\mysplit(φ,n+1) =
\begin{cases}
\unshunt(\mysplit(φ₁,n)) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂,n)
  & φ := φ₁ ∧ φ₂,\\

\unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n))
  & φ := φ₁ ∨ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(φ₂,n))
  & φ := φ₁ ⇒ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(φ₂,n)) ∧ \unshunt(φ₂ ⇒ \mysplit(φ₁,n))
  & φ := φ₁ ⇔ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂, n))
  & φ := ¬ (φ₁ ∧ φ₂),\\

\unshunt(\mysplit(¬ φ₁,n)) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂,n))
  & φ := ¬ (φ₁ ∨ φ₂),\\

\unshunt(\mysplit(φ₁,n)) ∧ \unshunt(φ₁ ⇒ \mysplit(¬ φ₂,n))
  & φ := ¬ (φ₁ ⇒ φ₂),\\

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂,n)) ∧ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁,n))
  & φ := ¬ (φ₁ ⇔ φ₂),\\

\unshunt(\mysplit(φ₁, n))
  & φ := ¬ ¬ φ₁,\\

⊤ & φ := ¬ ⊥,\\

⊥ & φ := ¬ ⊤.
\end{cases}
\end{equation}

@@

\begin{proof} We use induction on the structure of the formula
following the cases in the definition of \mysplit function.
We proceed to show some cases since the others are proved in a
similar way.\\[3mm]
(φ := φ₁ ∧ φ₂)
\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}_1$}
\UnaryInfC{Γ ⊢ φ₁}

@@

\AxiomC{$Γ ⊢ \mysplit(φ)$}
\RightLabel{By Def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\tt $∧$-proj$₂$}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{\tt $⇒$-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ φ₂$}

\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ φ₁ ∧ φ₂$}
\end{bprooftree}
\end{equation*}
where
\begin{equation*}
\mathcal{D}_1 =
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ)$}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁$}
\end{bprooftree}
\end{equation*}

(φ := φ₁ ⇒ φ₂)
\begin{equation*}
\begin{bprooftree}
\AxiomC{}
\RightLabel{\tt assume φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
\AxiomC{$Γ ⊢ \mysplit(φ)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{weaken φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{\tt ⇒-elim}
\BinaryInfC{$Γ , φ₁ ⊢ \mysplit(φ₂)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
\RightLabel{\tt ⇒-intro}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
\end{bprooftree}
\end{equation*}

(φ := ¬ (φ₁ ∨ φ₂))
\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}$}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}

@@

\AxiomC{$Γ ⊢ \mysplit(φ)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{\tt ⇒-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ ¬ φ₂$}
\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ ¬ φ₁ ∧ ¬ φ₂$}
\RightLabel{\tt ¬∧¬-to-¬∨}
\UnaryInfC{$Γ ⊢ ¬ (φ₁ ∨ φ₂)$}
\end{bprooftree}
\end{equation*}
where
\begin{equation*}
\mathcal{D} =
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
% (φ := ¬ (φ₁ ⇔ φ₂))
% \begin{equation*}
% \begin{bprooftree}

% % -- right

>>>>

The proof-reconstruction approach proposed here consists of a series
of steps similar to the workflow presented by Sultana in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is the proof
assistant, \Agda. We choose \Agda, but a proof assistant with
support of type theory and similar features like \Agda
(see a description of \Agda in section
\ref{sec:implementation}) can be used.
Following the diagram in Fig.
\ref{fig:proof-reconstruction-workflow}, the process can begin with
a problem in \CPL that we encode in \TPTP format to be sent as the
input of the \Metis prover.
The proof-search algorithms of \Metis in step No. 3 intend to find
a solution; if they success, \Metis replies a derivation in \TSTP
format of the proof, else \Metis may no stop and running forever or
replies the problem is counter feasible.

@@

In the rest of this section, we provide a formal description using
type theory to build definitions and theorems for the functions
necessary to emulate the \Metis' inference rules. We will write the
type judgments with ``$a : A$'' to denote that $a$ has type $A$.
For propositions, we will denote by $φ ≡ ψ$ the syntactic
equivalence.

@@

\subsection{Well-Structured Recursion}
\label{ssec:well-structured recursion}

In the process to emulate \Metis' inference rules in type theory,
We encountered many recursive algorithms with termination problems,
that is,when an algorithm could run forever or we can not guarantee
that it will stop after some point. Such a problem makes impossible
to translate the algorithm in type theory directly.
For that reason, we show a technique that will help us to define
new algorithms based on the previous ones to solve the problem
mentioned.

A recursive function is a \emph{well-structured recursion}, if the
size of the arguments with which it is defined decrease in
size for each call to the function in the definition~\cite{Bove2005}.
A well-structured function terminates and it can be translated
directly to type theory~\cite{Bove2002}.

Let us consider the following example from a inner function
in the \strip inference.

\begin{equation}
\label{eq:unshunt-definition}
\unshunt(φ) =
\begin{cases}
\unshunt((φ₁ ∧ φ₂) ⇒ φ₃)
  & φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\

\unshunt(φ₁ ⇒ φ₂) ∧ \unshunt(φ₁ ⇒ φ₃)
  & φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃),\\

@@

From the definition above, in the first case of the \unshunt
function, the size of the argument doesn't really decrease.
Actually, it maintains with the same length syntactically
(i.e., (φ₁ ∧ φ₂) ⇒ φ₃ compared with (φ₁ ⇒ (φ₂ ⇒ φ₃))).
Therefore, a termination checking will fail with \unshunt function.

To improve that definition for terminating, we will use the
technique proposed by Bove and Capettra in
\cite{Bove2005}, which in particular with functions that
transform propositions consists of:
take a general recursive function $f: \PROP \to \PROP$.
and a predicate $P: \PROP \to \abbre{Ind}$, where $\abbre{Ind}$
is a type with a structured inductive definition.

Then, we redefine the function as
$f: \PROP \to \abbre{Ind}\ \to \PROP$, and perform the recursion
using induction on the structure of the \abbre{Ind} type.

In our case, the \abbre{Nat} data type for
natural numbers will be the $\abbre{Ind}$ type.
We define \abbre{Nat} as usual (constructors \zero and \suc).
We also allow common arithmetic operations and numbers represented by
$0,1,2,\dots$ that stand for $\zero, \suc, \suc(\zero), \dots$.

Thus, we have a new function
$\unshunt : \PROP \to \Nat \to \PROP$
defined by structural recursion in its second argument in
Eq.~\ref{eq:unshunt-structured}.

\begin{align}
\label{eq:unshunt-structured}
\begin{split}
\unshunt(φ, \zero) &= φ\\
\unshunt(φ, \suc(n)) &=
\begin{cases}
\unshunt((φ₁ ∧ φ₂) ⇒ φ₃, n)
  & φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\unshunt(φ₁ ⇒ φ₂, n)  & φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).\\
∧\ \unshunt(φ₁ ⇒ φ₃, n)
\end{cases}
\end{split}
\end{align}

To feed the second argument of this new function, we will use
functions. For instance, we opted to use a complexity measure of
the formula that \unshunt function transforms.

@@

A \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
and presented in \cite{VanDalen1994} as well as \emph{the rank for a
proposition} is a function with type
$\PROP \to \Nat$ that assigns a natural number to a
proposition φ as follows.

\begin{equation*}
\label{eq:rank-definition}
\rank(φ)=
\begin{cases}
0 &\text{for atomic }φ, \\
\rank(φ₁) + c_{¬}  & φ ≡ \neg φ₁, \\
\max(\rank(φ₁),\rank(φ₂)) + c_{\square}
  & φ ≡ φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}

where $c_{¬}$ or $c_{\square}$ is a non-negative integer called the
complexity of the connective.

We have already solved the terminating problem for the \unshunt
function in Eq.~\ref{eq:unshunt-structured}, nevertheless we do not
what natural number should be feed that argument in
order to maintain the same outputs as the original \unshunt does in
Eq.~\ref{eq:unshunt-definition}.

Therefore, we define an appropriate complexity measure
$\unshunt_{cm}: \PROP → \NN$
for the second argument of the $\unshunt$ function in
Eq.~\ref{eq:unshunt-rank-definition}.

\begin{equation}
\label{eq:unshunt-rank-definition}
\unshunt_{cm}(φ)=
\begin{cases}
\unshunt_{cm}(φ₃) + 2 & φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\max(\unshunt_{cm}(φ₂),\unshunt_{cm}(φ₃)) + 1
  & φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃),\\
0 &\text{otherwise.}
\end{cases}
\end{equation}

Then, a curated version of the \unshunt functions is presented in Eq.
\ref{eq:unshunt-fixed}.

\begin{equation}
\label{eq:unshunt-fixed}
\unshunt(\varphi) := \unshunt(\varphi, \unshunt_{cm}(\varphi)).
\end{equation}

@@

\emph{Splitting a goal.}
The goal of a proof is processed by \Metis to generated new targets,
\Metis calls these new formulas \emph{subgoals}.
Then, \Metis proves each subgoal and with that it finishes the proof
of the goal.

Nevertheless, \Metis does not provide neither an explanation of this
process in its \TSTP derivations nor in its documentation.
We provide such a evidence that the process of creating subgoals,
and combining them in a final conjunction proves the original goal.

\newcommand{\thmunshunt}{\tt thm-$\unshunt$}
\begin{theorem}[\thmunshunt]
\label{thm:unshunt}
$Γ ⊢ \unshunt(φ)$, then $Γ ⊢ φ$.
\end{theorem}

\begin{proof} Use induction on the cases defined by the result of
$\unshunt$ in Eq.~\ref{eq:unshunt-structured}.
When $n = 0$, by definition, $Γ ⊢ φ$.
When $n = 1$, we apply the following theorem presented in
\cite{AgdaProp}

\begin{equation*}
\texttt{∧⇒-to-⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
\end{equation*}

For $n > 1$, we will use induction on the structure of second
argument.

(φ₁ ⇒ (φ₂ ⇒ φ₃))
\begin{equation*}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
\RightLabel{\tt thm-$\unshunt$}
\UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
\RightLabel{\tt ∧⇒-to-⇒⇒}
\UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).}
\end{bprooftree}
\end{equation*}

Second, using the following theorem from \cite{AgdaProp},
\begin{equation*}
\texttt{⇒∧-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{equation*}

(φ₁ ⇒ (φ₂ ∧ φ₃))
\begin{equation*}
\scalebox{0.95}{
($\mathcal{D}_1$)
\begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt(φ₁ ⇒ φ₂, n) ∧ \unshunt(φ₁ ⇒ φ₃, n)$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ φ₂, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ φ₂, n-1)$}
\RightLabel{\tt thm-$\unshunt$}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$,}
\end{bprooftree}
}
\end{equation*}

\begin{equation*}
(\mathcal{D}_2)
\begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt(φ₁ ⇒ φ₂) ∧ \unshunt(φ₁ ⇒ φ₃)$}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ φ₃, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ φ₃, n-1)$}
\RightLabel{\tt thm-$\unshunt$}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$,}
\end{bprooftree}
\end{equation*}

\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}_1$}
\AxiomC{$\mathcal{D}_2$}
\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
\RightLabel{\tt ⇒∧⇒-to-⇒∧}
\UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃).}
\end{bprooftree}
\end{equation*}
\end{proof}

Now, let us consider the \mysplit function that
yields the conjunction of subgoals from a proposition goal.

\begin{equation}
\label{eq:split-definition}
\mysplit(φ) =
\begin{cases}
\unshunt(\mysplit(φ₁))              & φ ≡ φ₁ ∧ φ₂,\\
∧\ \unshunt(φ₁ ⇒ \mysplit(φ₂))      &\\[2mm]

\unshunt(¬ φ₁ ⇒ \mysplit(φ₂))       & φ ≡ φ₁ ∨ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(φ₂))         & φ ≡ φ₁ ⇒ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(φ₂))         & φ ≡ φ₁ ⇔ φ₂,\\
∧\ \unshunt(φ₂ ⇒ \mysplit(φ₁))      &\\[2mm]

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂))       & φ ≡ ¬ (φ₁ ∧ φ₂),\\

\unshunt(\mysplit(¬ φ₁))            & φ ≡ ¬ (φ₁ ∨ φ₂),\\
∧\ \unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂))  &\\[2mm]

\unshunt(\mysplit(φ₁))              & φ ≡ ¬ (φ₁ ⇒ φ₂),\\
∧\ \unshunt(φ₁ ⇒ \mysplit(¬ φ₂))    &\\[2mm]

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂))       & φ ≡ ¬ (φ₁ ⇔ φ₂),\\
∧\ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁))    &\\[2mm]

\unshunt(\mysplit(φ₁))              & φ ≡ ¬ ¬ φ₁,\\
⊤                                   & φ ≡ ¬ ⊥,\\
⊥                                   & φ ≡ ¬ ⊤.
\end{cases}
\end{equation}

We reformulate \mysplit function to match with the
new definition after apply the process described above in
subsection~\ref{ssec:well-structured recursion}, since
in the recursion calls of \mysplit in Eq.~\ref{eq:split-definition},
the negations, $(¬ φ_{1,2})$ raise a terminating problem.

@@

\begin{proof} We now proceed by induction on the structure of the
formula φ in the definition of Eq.~\ref{eq:split-definition}.
For a complete proof, we refer the reader to \cite{AgdaMetis}.\\[3mm]

(φ ≡ φ₁ ∧ φ₂)
\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}_1$}
\UnaryInfC{Γ ⊢ φ₁}

@@

\AxiomC{$Γ ⊢ \mysplit(φ,n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\tt $∧$-proj$₂$}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{\tt $⇒$-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ φ₂$}

\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ φ₁ ∧ φ₂$.}
\end{bprooftree}
\end{equation*}
Where,
\begin{equation*}
(\mathcal{D}_1)\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ,n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁$.}
\end{bprooftree}
\end{equation*}

(φ ≡ φ₁ ⇒ φ₂)
\begin{equation*}
\begin{bprooftree}
\AxiomC{}
\RightLabel{\tt assume φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
\AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{weaken φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{\tt ⇒-elim}
\BinaryInfC{$Γ , φ₁ ⊢ \mysplit(φ₂,n-1)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
\RightLabel{\tt ⇒-intro}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$.}
\end{bprooftree}
\end{equation*}

(φ ≡ ¬ (φ₁ ∨ φ₂))
\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}$}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}

@@

\AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂,n-1)$}
\RightLabel{\tt ⇒-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ ¬ φ₂$}
\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ ¬ φ₁ ∧ ¬ φ₂$}
\RightLabel{\tt ¬∧¬-to-¬∨}
\UnaryInfC{$Γ ⊢ ¬ (φ₁ ∨ φ₂)$.}
\end{bprooftree}
\end{equation*}

Where,

\begin{equation*}
(\mathcal{D})
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁$.}
\end{bprooftree}
\end{equation*}
\end{proof}

\begin{theorem} If $s_i$ are the subgoals for the goal φ, and $\{ Γ ⊢ s_i \}_{i}$, then $Γ ⊢ φ$.
\end{theorem}

\begin{proof}
\begin{equation*}
\begin{bprooftree}
\AxiomC{$\left\{ Γ ⊢ s_i \right\}_{i}$}
\RightLabel{\tt ∧-intro}
\UnaryInfC{$Γ ⊢ \bigwedge_i s_i$}
